A Lyapunov Analysis of Momentum Methods \\ in Optimization

### 2016-11-08 Draft14.tex
Update Draft14.tex

\paragraph{General accelerated mirror descent} In Wibisono {\em et al.}~\cite{Acceleration}, the following Lyapunov function:
\begin{align}\label{Eq:E}
\E_t \,=\, e^{\beta_t} (f(X_t)-f(x^*)) + D_h\left(x^*, \, X_t + e^{-\alpha_t} \dot X_t\right),
\end{align}
was introduced to analyze the flow that forms the basis for accelerated mirror descent.
%  Notice, assuming we start from rest $\dot X_0 = 0$, the  Lyapunov property allows us to conclude,
 %\begin{equation*}
 %f(X_t) - f(x^\ast)\leq \frac{e^{\beta_0} (f(X_0) - f(x^\ast) )+ D_h(x^\ast, X_0)}{e^{\beta_t}}.
 %\end{equation*}
In Section~\ref{Sec:MomLyap}, we show how this Lyapunov function emerges from a general framework for analyzing momentum dynamics; furthermore, we show how to discretize the Lyapunov function and thereby obtain a tool which can be used to analyze a variety of algorithms that are discretizations of momentum dynamics.
%MJ:  I don't view this footnote as necessary; that derivation was in supplementary information \footnote{The same derivation of the Lyapunov function was added to~\cite[App. C]{Acceleration}}
%
%
%
%ACCC GRAD STRONG
\paragraph{Accelerated Gradient Descent (Strong Convexity)}
In this paper, we show that the following function 
   \begin{equation}\label{eq:general-sc-lyap}
  \E_t =  e^{\beta_t} \left(f(X_t) - f(x^\ast) + \mu D_h\Big(x^\ast ,X_t + e^{-\alpha_t} \dot X_t\Big)\right),
 \end{equation}
is a Lyapunov function for a family of momentum dynamics when $f$ is $\mu$-strongly convex ($D_f(y,x) \geq \mu D_h(y,x)$) and $e^{\alpha_t} = \dot \beta_t$.  In Section~\ref{Sec:StrongConv}, we derive this Lyapunov function and show how it can be used to demonstrate a linear rate of convergence for the family of accelerated gradient descent dynamics.  (For simplicity, we focus on the setting in which $h = \frac{1}{2}\|\cdot\|^2$ is Euclidean.) We also show that a particular discretization of the family of dynamics provides an algorithm for which a commensurate discretization of~\eqref{eq:general-sc-lyap} is a Lyapunov function.

@@

%
%
%
\subsubsection{Conditional Gradient Method} 
\paragraph{Continuous-time Estimate Sequence}
For the conditional gradient method, the following condition needs to hold:
\begin{align}\label{Eq:InEq4}
0 &\leq \int_0^t\dot \beta_t e^{\beta_t}  [f(X_s)+ \langle \nabla f(X_s), x - X_s \rangle] -  \frac{d}{ds}\left( e^{ \beta_s}  f(X_s)\right)\,ds.
\end{align}
 We can write this as
\begin{align*}
e^{ \beta_t}  f(X_t) \leq \int_0^t\dot \beta_t e^{\beta_t}  [f(X_s)+ \langle \nabla f(X_s), x - X_s \rangle]ds   + e^{\beta_0} f(X_0),
\end{align*}
and, from the derivation of the Lypaynov function, we conclude
\begin{align*}
e^{\beta_t} f(X_t) \leq  (e^{\beta_t} - e^{\beta_0}) f(x) + e^{\beta_0}f(X_0).
\end{align*}
Thus, $\{f(X_t),e^{\beta_t}\}$ is a continuous-time estimate sequence for this method. In Appendix~\ref{App:EstSeq}, we show the corresponding equivalence in discrete time.
%

>>>>

\paragraph{General accelerated mirror descent} In Wibisono {\em et al.}~\cite{Acceleration}, the following Lyapunov function:
\begin{align}\label{Eq:E}
\E_t \,=\, e^{\beta_t} (f(X_t)-f(x^*)) + D_h\left(x^*, \, X_t + e^{-\alpha_t} \dot X_t\right),
\end{align}
was introduced to analyze the flow that forms the basis for accelerated mirror descent.
%  Notice, assuming we start from rest $\dot X_0 = 0$, the  Lyapunov property allows us to conclude,
 %\begin{equation*}
 %f(X_t) - f(x^\ast)\leq \frac{e^{\beta_0} (f(X_0) - f(x^\ast) )+ D_h(x^\ast, X_0)}{e^{\beta_t}}.
 %\end{equation*}
In Section~\ref{Sec:MomLyap}, we show how this Lyapunov function emerges from a general framework for analyzing momentum dynamics; furthermore, we show how to discretize the Lyapunov function and thereby obtain a tool which can be used to analyze a variety of algorithms that are discretizations of momentum dynamics.
%MJ:  I don't view this footnote as necessary; that derivation was in supplementary information \footnote{The same derivation of the Lyapunov function was added to~\cite[App. C]{Acceleration}}
%
%
%
%ACCC GRAD STRONG
\paragraph{Accelerated gradient descent (strong convexity)}
In this paper, we show that the following function 
   \begin{equation}\label{eq:general-sc-lyap}
  \E_t =  e^{\beta_t} \left(f(X_t) - f(x^\ast) + \mu D_h\Big(x^\ast ,X_t + e^{-\alpha_t} \dot X_t\Big)\right),
 \end{equation}
is a Lyapunov function for a family of momentum dynamics when $f$ is $\mu$-strongly convex ($D_f(y,x) \geq \mu D_h(y,x)$) and $e^{\alpha_t} = \dot \beta_t$.  In Section~\ref{Sec:StrongConv}, we derive this Lyapunov function and show how it can be used to demonstrate a linear rate of convergence for the family of accelerated gradient descent dynamics.  (For simplicity, we focus on the setting in which $h = \frac{1}{2}\|\cdot\|^2$ is Euclidean.) We also show that a particular discretization of the family of dynamics provides an algorithm for which a commensurate discretization of~\eqref{eq:general-sc-lyap} is a Lyapunov function.

@@

%
%
%
\subsubsection{Conditional Gradient Method} 
\paragraph{Continuous-time estimate sequence}
For the conditional gradient method, the following condition needs to hold:
\begin{align}\label{Eq:InEq4}
0 &\leq \int_0^t\dot \beta_t e^{\beta_t}  [f(X_s)+ \langle \nabla f(X_s), x - X_s \rangle] -  \frac{d}{ds}\left( e^{ \beta_s}  f(X_s)\right)\,ds.
\end{align}
 We can write this as
\begin{align*}
e^{ \beta_t}  f(X_t) \leq \int_0^t\dot \beta_t e^{\beta_t}  [f(X_s)+ \langle \nabla f(X_s), x - X_s \rangle]ds   + e^{\beta_0} f(X_0),
\end{align*}
and, from the derivation of the Lypaynov function, we conclude
\begin{align*}
e^{\beta_t} f(X_t) \leq  (e^{\beta_t} - e^{\beta_0}) f(x) + e^{\beta_0}f(X_0).
\end{align*}
Thus, $\{f(X_t),e^{\beta_t}\}$ is a continuous-time estimate sequence for this method. In Appendix~\ref{App:EstSeq}, we show the corresponding equivalence in discrete time.
%
